\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[hidelinks]{hyperref}
\usepackage[a4paper, margin=3cm]{geometry}
\usepackage{parskip}

\title{On Meditation}
\date{07/06/26}

\begin{document}
\maketitle

\paragraph{v1}
In the Shivaratri Seminar in Brazil Uncle Eduardo Marino gave a presentation on attention and meditation during the Seminar kick-off. I found it extremely insightful and want to expand on it little bit. (The slides of his presentation can be found here.)

In the presentation he first defines the attention as the divine within us

- Definition of attention
- Patanjali’s Yoga Sutra and the 8 limbs of Yoga → Focus on the last 4
- How I learned to meditate with “how to get into meditation” with 4 main steps
- How those 4 steps are related to Patanjali
- Learning: Only the final step really requires Kundalini
That’s why other meditations also manage to have very low to no brainwaves during measurements
And that also means we usually fail at the initial steps if we don’t get into meditation/samadhi at all (That’s basically the message I want to give: Take to meditation somewhat methodically and go step by step).

New Idea
- Uncle’s presentation including the yoga sutras inspired me to reflect on how i get into meditation
- Based on Mother’s ‘How to get into meditation’ I do something like the following

---

In the Shivaratri Seminar in Brazil Uncle Eduardo Marino gave a presentation on attention and meditation during the Seminar kick-off. I found it extremely insightful and got inspired to reflect on my own approach to meditation. (Check out the slides of his presentation here.)

Firstly, he talks about the divine within us - which is the attention - and that this attention is the spectator of a screen that our mind uses to project its images on. Depending on the state of our awareness our attention gets dragged towards these projections of the mind, which creates a thought process and takes us into the realm of the mind, i.e. the realm of illusion. 

Further, the presentation pivots to Patanjali’s Yoga Sutras (i.e. the 8 limbs of Yoga) and determines the definition of Yoga as the cessation of the movement of the attention. The 8 limbs describe the path towards the Yoga as follows:

1. Yama     (Mastering  External Righteousness)
2. Nyama (Mastering  Inner Righteousness)
3. Pranayama (Mastering the Breath Process)
4. Asana   ( Mastering Yoga Postures)
5. Pratyahara (Moving the Attention Inwards)
6. Dharana  (Concentrating the Attention)
7. Dhyana    (Ceasing Thoughts - Meditation)
8. Samadhi  (Thoughtless Awareness - Nirvichara)

Hereby, the most important are the final 4 limbs that describe the subtle parts of Yoga, which are also most relevant in Sahaja Yoga.

When I saw that, there was an immediate feeling of association and it reminded me of how I learned to go into meditation. From a guided meditation by Shri Mataji from 1979, I derived the following method.

First, we put the attention on the heart and invite Shri Mataji’s lotus feet into our heart. After that we take the attention to our Mooladhara Chakra and invite Shri Ganesha to receive the authorization for our Kundalini to rise. The next step involves taking the attention to the mind and its thoughts and thereby focusing the attention by gathering it within. The question: “Where is my attention?” may serve as a good aid. During this process the waves of the attention start matching the waves of the thoughts and slowly the thought process recedes. This is also a good moment to utilize the *Nirvichara Mantra* to fully enter a state of thoughtlessness. Now, the thought process is ceased, however, that doesn’t mean that the Agya is opened fully. With the help of the *Mahatahankara Mantra* we can bring down the Ego and give space to our Kundalini to pass through the tiny gate. It is for this moment that we do all the previous steps - just like prerequisites - to offer our Kundalini an Agya that is somewhat stable and steady, so that a passage is offered to her for passing. If our attention were still dragged around by the mind, pulled by the Ego or Super-Ego, the passage for our Kundalini would be impossible. Finally, the Kundalini reaches Sahastrara and we start going deeper into our being and into the connection with the divine. At this point we can take the *Sahastrara Mantra* and tie our Kundalini up there how we do during *Bandhan.*

This process can be divided into 4 steps:

- Taking the attention to the heart
- Watching the thoughts
- Increasing the gap between thoughts
- Opening the Agya and settling the attention on Sahastrara

and if we compare this to the Yoga Sutras, we see the analogy clearly:

| Pratyahara | Moving the attention inward | Taking the attention on the heart |
| --- | --- | --- |
| Dharana | Concentrating the attention | Watching the thoughts |
| Dhyana | Ceasing thoughts | Increasing the gap between thoughts |
| Samadhi | Thoughtless awareness | Opening Agya and connecting with divine in Sahatrara |

One question I often had was what role the method of Sahaja Yoga played in this process and how the Kundalini is related to the Yoga Sutras. Do I need an awakened Kundalini to be thoughtless? This is at least something that Sahaja Yogis often answer with: “yes.” If that were so, we should only see brainwave measurements (e.g. with an EEG) indicating a thoughtless state for people who have an awakened Kundalini. However, that is not the case. There are numerous Buddhists and other meditators, who achieve a state of thoughtlessness - or *Dhyana*, as Patanjali would say. To stop the thought process can be achieved by effort and by controlling the mind. Where the Kundalini becomes necessary is for passing through the Agya - or for entering *Samadhi* in Patanjali’s words. That also means, even an awakened Kundalini cannot fully replace the prerequisite steps that are necessary to open the Agya. A mind out of control can impossibly be risen to a higher state. It is that part that we have to accomplish before the Kundalini can take over. I like to think of an analogy of a small boy and its mother. Say, the boy fell while playing and wounded his knee and comes crying to his mother for help. Now, the mother knows exactly what to do. She fetches a plaster and some disinfectant to apply on the knee. In order to be helped the boy is required to sit down and be still for the time being. Only once his knee is stationary she can apply the disinfectant and the plaster. Imagine now the boy keeps running around in the room and doesn’t calm down even though he might be in pain. What can the mother do? Even if she tries her best and runs alongside the boy, anticipating his movements, and then applying disinfectant and a plaster, the chances for success are very low. If we apply this to ourselves, where the mind is the boy and the Kundalini the mother, it is of crucial importance to calm the mind, make it stationary, seize its movement as much as possible in order for the Kundalini to begin its healing process. If we cannot accomplish this little task we cannot be helped. This also sheds light on the paradox of effort in the context of meditation, which is supposed to be completely effortless. In the words of the analogy the answer would be: Our effort should be to become stationary, while the healing process should be left completely to the divine power. Just like the boy better just watch the process instead of asking too many questions about the plaster, or giving advice on how to apply the disinfectant properly.

To conclude I want to frame this as an encouragement for all of us. There is a lot we can do to improve our meditation, especially when it comes to calming our mind. We don’t have to leave it to chance, the weather, or the current planetary constellation whether we go thoughtless or not. We can simply practice the control over our mind, so that our Kundalini can perform her miracle.

% Brainstorm v2
% When we talk about meditation, what do we actually mean?
% Patanjali has defined 4 relevant steps towards Yoga, i.e. Pratyahara, Dharana, Dhyana, and Samadhi.
% Now, according to Western philosophy Pratyahara could already define meditation as the inward movement of the attention, or as contemplation. Just like in the term "premeditate".
% In \textit{Mindfulnes} meditation is probably Dharana - focusing the attention, while in Buddhism it would be Dhyana that best defines meditation, namely the cessation of thought.
% Lastly, in Sahaja Yoga we talk of Samadhi when we say meditation, as that is the state of thoughtless awareness, which is enabled by the crossing of the Kundalini through the Agya.
%
% Furthermore, when we talk about meditation we have to also explain what attention really is.
% When we talk about meditation, what do we actually mean?
% Patanjali has defined 4 relevant steps towards Yoga, i.e. Pratyahara, Dharana, Dhyana, and Samadhi.
% Now, according to Western philosophy Pratyahara could already define meditation as the inward movement of the attention, or as contemplation. Just like in the term "premeditate".
% In \textit{Mindfulnes} meditation is probably Dharana - focusing the attention, while in Buddhism it would be Dhyana that best defines meditation, namely the cessation of thought.
% Lastly, in Sahaja Yoga we talk of Samadhi when we say meditation, as that is the state of thoughtless awareness, which is enabled by the crossing of the Kundalini through the Agya.
%
% Furthermore, when we talk about meditation we have to also explain what attention really is.

% Take 2
\paragraph{v2}

% Thesis
Most Sahaja Yogis don't actually know how to meditate.

% Intro
When I attend SY courses it constantly happens that one of the speakers eventually gets the question: "How long will it take to reach this thoughtless awareness?"
  and answers something like: "Puh, haha. That can take a long time. You know I've been meditating for almost 30 years now and I'm just scratching the surface."
Of course, there is a natural humility that wants to be expressed.
However, often enough the vibe is rather: "I'm struggling to get there since 30 years, which creates doubt in myself. At the same time feel superior to you.
  That's why I actually don't want you to achieve it, or I believe you won't achieve before me anyway and that's why I won't give you any hope."
And that bothers me. A lot.
Some years ago I came across a guided meditation by Shri Mataji from 1979, titled \textit{How to get into meditation}.
Upon following her instructions I was completely struck and shook - I have just learned how to actually meditate.
After 5 years of seriously practicing SY and a whole lifetime in the SY community, it was a random and not particularly popular video of around 20 minutes
  that would be the key for my spiritual journey.
Another thing that I noticed more and more was how the typical guided meditations among Sahaja Yogis resembled very little to what SM covers in that video.
We most commonly first balance our channels and then work ourselves up along the Chakras while chanting different Mantras until we reach the Sahastrara, where
  we then remain in silence for some time.
We then say things like: "And now let's keep the attention on Sahastrara and feel the vibrations."
In the guided meditation the sequence is more like follows.
First, the attention goes to the heart, where the lotus feet of the Guru are placed in the heart.
Then, there is space to balance the channels or invoke Shri Ganesha to support the rising of the Kundalini.
After that the attention goes towards the mind with the goal to steady the attention.
For that the question: "Where is my attention?" gives the guidance.
Once the attention is stabilized and held in control, the \textit{ Nirvichara } Mantra is chanted to create the complete cessation of thought, i.e. thoughtlessness.
The opening of Agya follows only after that, namely by bringing down the Ego.
Either with the \textit{ Mahat Ahankara } Mantra, or literally by bring the Ego down using the right hand that lifts the left side and lowers the right.
Now, the real meditation starts as the Kundalini passes through the Agya and touches the Sahastrara, enlightening the being with the light of the spirit and
  establishing a connection, or union, with the collective unconscious.
What is worth noting, is that the process is very methodical, as we will see in a moment.
For years until now this had become the method for me to get into meditation on a daily basis.
Always the same.
Morning and evening.
And I must say, I don't understand why almost no guided meditations in Sahaja Yoga follow this procedure, and neither did I understand the theory behind the method.

That changed this year in January, when we attended the Brazil Tour and the subsequent Shivaratri seminar in Brasilia.
In the inauguration, or the welcome ceremony, of the seminar, Uncle Eduardo Marino gave a presentation on attention and how to meditate.
He would present his understanding of attention and its importance for the meditation compiled from different speeches of SM.
Additionally, he would take Patanajali's \textit{ Yoga Sutras } and the 8 limbs of Yoga to support the message, and lasty he guided a meditation
  that would put in practice the presented theory.
For me it was an absolute excitement, as I had finally attended a Sahaj event, where the guided meditation resembled very well the procedure from
  that 1979 guided meditation by SM.
Later, I also started understanding the reason why this method worked so well.

As, Uncle Eduardo would point out in his presentation there are 8 limbs in the \textit{Yoga Sutras}, where the first 4 represent outward practice, like
  \textit{Pranayama}, or \textit{Asana}, while the final 4 represent the inward practice.
They are \textit{Pratyahara}, \textit{Dharana}, \textit{Dhyana}, and finally, \textit{Samadhi}.
Pratyahara means the drawing of the attention towards the inside.
Dharana describes the focusing of the attention, and Dhyana the complete cessation of thought.
And finally, Samadhi means the state of awareness, where Yoga happens, or when we are in Yoga, i.e. union.

Let's now look at Shri Mataji's guided meditation and see how it connects.
Firstly, the attention on the heart, with the purpose of inviting the Guru into our heart - that is the Pratyahara.
We pull our attention inward completely and look for the divinity within.
After that, the question: "Where is my attention?" represents the Dharana.
Here we collect and steady our attention that is in movement due to mental projections from the mind that drag our attention towards it.
Once the attention is steady the Dhyana begins that consists of increasing the gap between the thoughts until a complete stillness is achieved.
For this SM recommends the Nirvichara Mantra.
And finally, the Mahat Ahankara Mantra represents the beginning of the final phase - the Samadhi - where a complete oneness is achieved.

Therefore, when we talk of thoughtless awareness in Sahaja Yoga we actually mean Samadhi.
That is a very importatnt distinction because only for this final stage of Yoga the awakening of the Kundalini is mandatory.
All the previous steps, from Pratyahara to Dhyana don't actually require Kundalini awakening.
If that weren't the case we would find that only Kundalini awakening leads to measurable thoughtlessness via EEGs, but that is not the case.
Buddhists and practitioners of other types of meditation achieve measurable stillness in their brainwaves just as well, and we know that the vast
  majority didn't receive their Self-Realization, or Kundalini awakening.
And also, if Sahaja Yoga's thoughtless awareness weren't the equivalent of Samadhi in the Yoga Sutras, then what's so special about Sahaja Yoga?
It would just be another meditation technique.
And that is clear to all of us - Sahaja Yoga must be something extraordinary.

Okay.
So, what do I want to say with this?
Getting into meditation is not someting "up in the air" or vague, as we often make it seem
  in our courses.
The answer is not only: "You just have to desire it more and ask your Kundalini to give you this state."
In reality it is a practice and it requires willpower and individual effort.
Until Dhyana, it's mostly in our hand.
To turn the attention inward, and focus it, is absolutely not Kundalini's job.
It is the self-respect we have to show to make ourself worthy for our Kundalini to rise in the
  first place.
From Dhyana to Samadhi, on the other hand, we are completely left to her will to bring us anywhere.
Opening the Agya cannot be achieved by our individual effort nor by our spiritual prowess.
Only the grace of the Kundalini allows the passage through the narrow gate and towards the union.

Getting into meditation should not be a stochastic event, that depending on the weather,
  or the alignment of the stars, it may happen to us or not.
On the contrary it should be the strongest tool of a Sahaja Yogi, a reliable process to
  align ourselves with the Paramchaitanya - whenever we desire to.
For that we just need more method.
We need to find out what disturbs our attention, where its movement goes and how to control it.
We also need to practice focus and honesty with ourselves.
It should be transparent to us what desires lure inside and how our Ego tries to escape exposure.

We can also illustrate this with an analogy.
Say, we are a small child and there is our mother that is taking care of us.
She represents our Kundalini, who protects us, nurtures us, and looks after us.
Imagine the child playing outside, running around, jumping over obstacles, and being fully immersed
  in its imagination - just like us how we manouver throught life.
Eventually, something goes wrong and the child falls badly and hurts itself.
The knee is wounded and is bleeding.
Immediately, it cries for its mother and asks for help.
Just like us when we face a difficult siutation in life and we need our mother to help us.
The child runs to its mother, crying loudly, and with great concern the mother starts tending
  to the child.
First, calming it down, then sitting it down to look at the wound.
After a quick diagnosis she will then prepare her treatment that consists of disinfecing the
  wound and applying a bandage to stop the bleeding.
If we imagine ourselves again, it could be that we had very bad day and with full despair we
  turn to Shri Mataji and ask her to relieve our innper pain by taking away all the negativity
  with us that causes our suffering.
Ideally, the child would stay calm and the mother slowly and carefully executes all parts of
  the treatment with great care for the wounded knee.
However, now imagine the child would continue running around in the house.
Maybe crying, maybe screaming for help.
How would the mother be able to help the child?
She tries to run behind the child, but doesn't manage to apply the disinfectant properly.
Further, the bandage she could manage to put on wouldn't hold properly because of the constant
  movement of the child.
In other words, it's a mess.
Now, think of the child as our mind.
How can the Kundalini tend to our problems if we cannot steady our attention.
If due to our lack of discipline our attention is wobbly and makes our Agya Chakra move
  in high speed, how will she find a passage through that Chakra?
How will she manage to ease our pain by subdueing our Ego and Super-Ego, if we don't give her
  the necessary space to treat us, i.e. if we keep running around or moving like the child?
She cannot.
Not because of her inability, but because of her respect of our free will.
She doesn't want to force anything on us.
We have to be mature enough to prepare ourselves accordingly so that she can treat us with all
  the love, affection and care that she wishes to.

That is the reason why Pratyahara, Dharana and Dhyana are for us to provide to our Kundalini,
  so that she can take us to Samadhi, to the Sahastrara.

In conclusion, thoughtless awareness in Sahaja Yoga corresponds to the Samadhi state as described in
  the Yoga Sutras by Patanajali and is preceded by 3 other steps that together build a simple
  method for getting into mediation.                         
This method helps us to identify at which stages we might have difficulties and should lead
  us toward a more successfull, and consistent outcome of our meditations.


% ==============================================================
% V3 — WORKING DRAFT OUTLINE
% Audience: Sahaja Yoga community
% Form: Argumentative essay with personal essay backbone
% ==============================================================
% [ADD: Write one arguable thesis sentence here before drafting further.
%  From our conversation the candidate is:
%  "Meditation has a learnable method — whether or not anyone teaches it to you properly,
%  finding that method is your responsibility."]
% ==============================================================

\paragraph{v3}

% ──────────────────────────────────────────────────────────────
% SECTION I: OPENING — The culture of passivity
% Purpose: Establish the problem without attacking anyone.
%          The reader should recognise the culture, not feel accused.
%          Close with one sentence that cracks the frame open.
% ──────────────────────────────────────────────────────────────

When I attend SY courses it constantly happens that one of the speakers eventually gets the question: ``How long will it take to reach this thoughtless awareness?''
and answers something like: ``Puh, haha. That can take a long time. You know I've been meditating for almost 30 years now and I'm just scratching the surface.''
Of course, there is a natural humility that wants to be expressed.

% [ADAPT — v2: The sentence below frames the speaker's intent as malicious.
%  Rewrite so it names the belief the speaker genuinely holds — one that many of us have absorbed without questioning.
%  The target: the problem is the culture, not the person.]
However, often enough the vibe is rather: ``I'm struggling to get there since 30 years, which creates doubt in myself.
At the same time feel superior to you.
That's why I actually don't want you to achieve it, or I believe you won't achieve before me anyway and that's why I won't give you any hope.''

% [ADAPT — v2: "And that bothers me. A lot." — too reactive for the new framing.
%  Replace with a sentence that names this as a cultural assumption we have all absorbed, not just a flaw in one speaker.]
And that bothers me.
A lot.

% [ADD: One declarative sentence that cracks the frame open. States that the assumption is wrong
%  and that the reader can know better. Should be a statement, not a question.]


% ──────────────────────────────────────────────────────────────
% SECTION II: THE DISCOVERY — The 1979 video
% Purpose: Establish authority through personal testimony.
%          Deliver the before/after moment. Frame yourself as student, not critic.
% ──────────────────────────────────────────────────────────────

Some years ago I came across a guided meditation by Shri Mataji from 1979, titled \textit{How to get into meditation}.
Upon following her instructions I was completely struck and shook --- I have just learned how to actually meditate.
After 5 years of seriously practicing SY and a whole lifetime in the SY community, it was a random and not particularly popular video of around 20 minutes
that would be the key for my spiritual journey.

% [ADD: One or two sentences on the before-state — what did meditation feel like before this video?
%  What were you doing instead? This is the "before" that makes the "after" meaningful.
%  Without it, the reader cannot feel the stakes of the discovery.]

Another thing that I noticed more and more was how the typical guided meditations among Sahaja Yogis resembled very little to what SM covers in that video.
We most commonly first balance our channels and then work ourselves up along the Chakras while chanting different Mantras until we reach the Sahastrara, where
we then remain in silence for some time.
We then say things like: ``And now let's keep the attention on Sahastrara and feel the vibrations.''

In the guided meditation the sequence is more like follows.
First, the attention goes to the heart, where the lotus feet of the Guru are placed in the heart.
Then, there is space to balance the channels or invoke Shri Ganesha to support the rising of the Kundalini.
After that the attention goes towards the mind with the goal to steady the attention.
For that the question: ``Where is my attention?'' gives the guidance.
Once the attention is stabilized and held in control, the \textit{Nirvichara} Mantra is chanted to create the complete cessation of thought, i.e.\ thoughtlessness.
The opening of Agya follows only after that, namely by bringing down the Ego.
Either with the \textit{Mahat Ahankara} Mantra, or literally by bringing the Ego down using the right hand that lifts the left side and lowers the right.
Now, the real meditation starts as the Kundalini passes through the Agya and touches the Sahastrara, enlightening the being with the light of the spirit and
establishing a connection, or union, with the collective unconscious.

For years until now this had become the method for me to get into meditation on a daily basis.
Always the same.
Morning and evening.
And I must say, I don't understand why almost no guided meditations in Sahaja Yoga follow this procedure, and neither did I understand the theory behind the method.


% ──────────────────────────────────────────────────────────────
% SECTION III: THE FRAMEWORK — Patanjali's four stages
% Purpose: Give the reader vocabulary before you use it.
%          Ground the method in something ancient and documented.
%          Stay in the role of student — the seminar is where you learned this.
% ──────────────────────────────────────────────────────────────

That changed this year in January, when we attended the Brazil Tour and the subsequent Shivaratri seminar in Brasilia.
In the inauguration, or the welcome ceremony, of the seminar, Uncle Eduardo Marino gave a presentation on attention and how to meditate.
He would present his understanding of attention and its importance for the meditation compiled from different speeches of SM.
Additionally, he would take Patanjali's \textit{Yoga Sutras} and the 8 limbs of Yoga to support the message, and lastly he guided a meditation
that would put in practice the presented theory.
For me it was an absolute excitement, as I had finally attended a Sahaj event, where the guided meditation resembled very well the procedure from
that 1979 guided meditation by SM.
Later, I also started understanding the reason why this method worked so well.

% [OPTIONAL — v1: Consider including the definition of attention before presenting Patanjali,
%  drawn from Uncle Eduardo's framing:
%  "Firstly, he talks about the divine within us --- which is the attention --- and that this attention
%  is the spectator of a screen that our mind uses to project its images on. Depending on the state of
%  our awareness our attention gets dragged towards these projections of the mind, which creates a thought
%  process and takes us into the realm of the mind, i.e.\ the realm of illusion."
%  Decision: include or skip. Adds conceptual depth but may slow the pace here.]

As Uncle Eduardo would point out in his presentation there are 8 limbs in the \textit{Yoga Sutras}, where the first 4 represent outward practice, like
\textit{Pranayama}, or \textit{Asana}, while the final 4 represent the inward practice.

% [NOTE — v1: Consider presenting the four inward limbs as a numbered list rather than prose,
%  so the reader has a reference to return to:
%  1.\ Pratyahara (Moving the Attention Inwards)
%  2.\ Dharana (Concentrating the Attention)
%  3.\ Dhyana (Ceasing Thoughts --- Meditation)
%  4.\ Samadhi (Thoughtless Awareness --- Nirvichara)
%  The prose sentences below (v2) are the alternative — decide which register fits better.]

They are \textit{Pratyahara}, \textit{Dharana}, \textit{Dhyana}, and finally, \textit{Samadhi}.
\textit{Pratyahara} means the drawing of the attention towards the inside.
\textit{Dharana} describes the focusing of the attention, and \textit{Dhyana} the complete cessation of thought.
And finally, \textit{Samadhi} means the state of awareness, where Yoga happens, or when we are in Yoga, i.e.\ union.


% ──────────────────────────────────────────────────────────────
% SECTION IV: THE MAPPING — SM's method is Patanjali's method
% Purpose: Deliver the structural aha. Should feel like confirmation,
%          not instruction. Keep this section tight.
% ──────────────────────────────────────────────────────────────

% [ADAPT — v2: "Let's now look at Shri Mataji's guided meditation and see how it connects."
%  Slightly informal. Consider whether to keep or replace with a tighter transition sentence.]
Let's now look at Shri Mataji's guided meditation and see how it connects.

Firstly, the attention on the heart, with the purpose of inviting the Guru into our heart --- that is the \textit{Pratyahara}.
We pull our attention inward completely and look for the divinity within.
After that, the question: ``Where is my attention?'' represents the \textit{Dharana}.
Here we collect and steady our attention that is in movement due to mental projections from the mind that drag our attention towards it.
Once the attention is steady the \textit{Dhyana} begins that consists of increasing the gap between the thoughts until a complete stillness is achieved.
For this SM recommends the \textit{Nirvichara} Mantra.
And finally, the \textit{Mahat Ahankara} Mantra represents the beginning of the final phase --- the \textit{Samadhi} --- where a complete oneness is achieved.

% [NOTE — v1: Include the table from v1 here, converted to a LaTeX tabular environment.
%  The markdown syntax will not compile. Adapt as follows and uncomment when ready:]
%
% \begin{tabular}{lll}
% \textit{Pratyahara} & Moving the attention inward  & Taking the attention to the heart           \\
% \textit{Dharana}    & Concentrating the attention  & Watching the thoughts                       \\
% \textit{Dhyana}     & Ceasing thoughts             & Increasing the gap between thoughts         \\
% \textit{Samadhi}    & Thoughtless awareness        & Opening Agya and connecting with the divine \\
% \end{tabular}


% ──────────────────────────────────────────────────────────────
% SECTION V: THE KEY INSIGHT — Samadhi is where Kundalini becomes necessary
% Purpose: The intellectual peak of the essay. Reframes the entire
%          conversation about Kundalini and individual effort.
%          This is the most original claim — give it room.
% ──────────────────────────────────────────────────────────────

% [NOTE — v1: Consider opening with the framing question before stating the insight — it mirrors
%  the reader's likely doubt and makes the answer land harder:
%  "One question I often had was what role the method of Sahaja Yoga played in this process and how the
%  Kundalini is related to the Yoga Sutras. Do I need an awakened Kundalini to be thoughtless?
%  This is at least something that Sahaja Yogis often answer with: `yes.'"]

Therefore, when we talk of thoughtless awareness in Sahaja Yoga we actually mean \textit{Samadhi}.
That is a very important distinction because only for this final stage of Yoga the awakening of the Kundalini is mandatory.
All the previous steps, from \textit{Pratyahara} to \textit{Dhyana}, don't actually require Kundalini awakening.
If that weren't the case we would find that only Kundalini awakening leads to measurable thoughtlessness via EEGs, but that is not the case.
Buddhists and practitioners of other types of meditation achieve measurable stillness in their brainwaves just as well, and we know that the vast
majority didn't receive their Self-Realization, or Kundalini awakening.

And also, if Sahaja Yoga's thoughtless awareness weren't the equivalent of \textit{Samadhi} in the Yoga Sutras, then what's so special about Sahaja Yoga?
It would just be another meditation technique.
And that is clear to all of us --- Sahaja Yoga must be something extraordinary.

% [NOTE — v1: This sentence adds precision and closes the insight cleanly:
%  "Where the Kundalini becomes necessary is for passing through the Agya ---
%  or for entering Samadhi in Patanjali's words."]


% ──────────────────────────────────────────────────────────────
% SECTION VI: THE CHALLENGE — The first three stages are yours
% Purpose: Deliver the practical demand. Frame as agency, not criticism.
%          This is where the essay earns its right to challenge the reader.
% ──────────────────────────────────────────────────────────────

% [ADAPT — v2: "Okay. So, what do I want to say with this?" — too conversational, remove.
%  Open this section directly with the statement below.]

Getting into meditation is not something ``up in the air'' or vague, as we often make it seem in our courses.
The answer is not only: ``You just have to desire it more and ask your Kundalini to give you this state.''
In reality it is a practice and it requires willpower and individual effort.

Until \textit{Dhyana}, it's mostly in our hand.
To turn the attention inward, and focus it, is absolutely not Kundalini's job.
It is the self-respect we have to show to make ourselves worthy for our Kundalini to rise in the first place.
From \textit{Dhyana} to \textit{Samadhi}, on the other hand, we are completely left to her will to bring us anywhere.
Opening the Agya cannot be achieved by our individual effort nor by our spiritual prowess.
Only the grace of the Kundalini allows the passage through the narrow gate and towards the union.

Getting into meditation should not be a stochastic event, that depending on the weather,
or the alignment of the stars, it may happen to us or not.
On the contrary it should be the strongest tool of a Sahaja Yogi, a reliable process to
align ourselves with the Paramchaitanya --- whenever we desire to.
For that we just need more method.
We need to find out what disturbs our attention, where its movement goes and how to control it.
We also need to practice focus and honesty with ourselves.
It should be transparent to us what desires lure inside and how our Ego tries to escape exposure.

% [ADD: A sentence that directly names the accountability — that this is not someone else's job to teach you.
%  Without it, the section lists what to do but stops short of naming the responsibility squarely.
%  This is the essay's most direct challenge to the reader.]


% ──────────────────────────────────────────────────────────────
% SECTION VII: THE ANALOGY — Mother and child
% Purpose: Illustrate the division of labour between effort and grace.
%          Arrives AFTER the argument — it clarifies, it does not conclude.
%          The v2 version is too long. Comments below mark what to cut.
% ──────────────────────────────────────────────────────────────

Say, we are a small child and there is our mother that is taking care of us.
She represents our Kundalini, who protects us, nurtures us, and looks after us.

% [CONDENSE — v2: The setup (child playing, falls, runs to mother, mother diagnoses) runs too long.
%  Keep only the essential beats. Target: two sentences maximum for the setup before the treatment begins.
%  The paragraph that begins "If we imagine ourselves again" (v2) should be cut entirely —
%  it breaks the analogy before it resolves.]
Imagine the child playing outside, running around, jumping over obstacles, and being fully immersed
in its imagination --- just like us how we manoeuvre through life.
Eventually, something goes wrong and the child falls badly and hurts itself.
The child runs to its mother, crying loudly, and with great concern the mother starts tending to the child.
First, calming it down, then sitting it down to look at the wound.
After a quick diagnosis she will then prepare her treatment that consists of disinfecting the wound and applying a bandage to stop the bleeding.

Ideally, the child would stay calm and the mother slowly and carefully executes all parts of
the treatment with great care for the wounded knee.
However, now imagine the child would continue running around in the house.
How would the mother be able to help the child?
She tries to run behind the child, but doesn't manage to apply the disinfectant properly.
Further, the bandage she could manage to put on wouldn't hold properly because of the constant movement of the child.

% [CUT — v2: The questions beginning "If due to our lack of discipline..." and
%  "How will she manage to ease our pain..." are redundant once the analogy resolves.
%  Move directly to the application below.]

Now, think of the child as our mind.
How can the Kundalini tend to our problems if we cannot steady our attention.
She cannot.
Not because of her inability, but because of her respect of our free will.
We have to be mature enough to prepare ourselves accordingly so that she can treat us with all
the love, affection and care that she wishes to.

That is the reason why \textit{Pratyahara}, \textit{Dharana} and \textit{Dhyana} are for us to provide to our Kundalini,
so that she can take us to \textit{Samadhi}, to the Sahastrara.

% [NOTE — v1: Consider closing the analogy with this sentence — it adds a light touch and lands with an image:
%  "Just like the boy better just watch the process instead of asking too many questions about the plaster,
%  or giving advice on how to apply the disinfectant properly."]


% ──────────────────────────────────────────────────────────────
% SECTION VIII: CLOSE — Find your method
% Purpose: End on encouragement, not lecture. Not a summary.
%          Leave the reader with something to do.
% ──────────────────────────────────────────────────────────────

% [ADAPT — v1: "To conclude I want to frame this as an encouragement for all of us." — too formal and meta.
%  The close should just be the encouragement, not announce it. Drop this sentence and open with the next.]
There is a lot we can do to improve our meditation, especially when it comes to calming our mind.
We don't have to leave it to chance, the weather, or the current planetary constellation whether we go thoughtless or not.
We can simply practice the control over our mind, so that our Kundalini can perform her miracle.

% [ADD: A sentence that speaks to the reader's individual path — your method is an example, not a prescription.
%  This is the sentence that holds both the humility and the challenge of the essay together.
%  The reader should leave feeling capable, not instructed.]

% [REMOVE — v2: The final two sentences (beginning "In conclusion, thoughtless awareness...")
%  summarise the argument. Do not repeat the argument in the close.
%  If there is a forward-looking thought in them, extract it; otherwise cut entirely.]

\end{document}

