\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[a4paper, margin=3cm]{geometry}
\usepackage{parskip}

\date{\today}

\begin{document}

\raggedright

{\small\the\year}

\bigskip

Dear Mother,

\bigskip

With great excitement I wish to present to you a new project that we are pursuing.
It is a writing setup that shall enable us to write pieces for the inspiration of our fellow members of society.

There is a grave pain inside me that we don't have the tools and the material ready to make people understand the greatness of your Sahaja Yoga.
Of course, it is impossible to convince an egoistic person to seek God.
Nevertheless, your incarnation and the knowledge of Kundalini and Self-Realisation are just too fundamentally true to not explain them on all levels of understanding.
My goal is that every sane mind can admit the fact that your work of Sahaja Yoga is extremely relevant to every human being.
For that I plan to pick out different current topics of relevance and connect them to Sahaja Yoga, or more precisely, expose that modern science cannot explain it properly.
For example, take the modern trend of atheism.
How can it be that for thousands of years human life revolved around God and since a bit more than 100 years we pretend it was all a hoax and God never actually existed.
That is not rational!
Or how the limbic area can be excited by many activities or substances, but none of them manage to excite it in a sustainable way, like the Kundalini does.
Why should we have a limbic area in the first place if there is seemingly no use for it?

With questions like this I want to show that we are missing something - something so central to our life that we should start looking for.
My hope is that more people start seeking through these pieces and start accepting that spirituality is not a vague thing, but a deep science itself,
and ultimately become able to appreciate your creation, your incarnation, and your relentless work.

With my most humble attitude I seek for your blessings and for Shri Ganesha's approval.
Please guide my writing and let every piece and every word carry the love and benevolence of the spirit.

\bigskip

In deepest gratitude and commitment to your vision, \\
Your son

\end{document}
